---
title: "HWK11"
author: "Misael Santana"
date: "December 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## 1)	Go through the electric equipment example. 
### a)	Plot the time series. Can you identify seasonal fluctuations and/or a trend? 
```{r , echo=FALSE }
library(fpp) #fpp package must be installed first
data(ukcars)
plot(ukcars)
str(ukcars)
```

I can see seasonal flictuation inside each year and a trend to increase the cars

### b)	Use a classical decomposition to calculate the trend-cycle and seasonal indices. 
```{r }
fitd <- decompose(ukcars)
plot(fitd)

```

### c)	Do the results support the graphical interpretation from part (a)? 
Yes I can see a variatoin along each year and a global trend

### d)	Compute and plot the seasonally adjusted data. 

```{r }
eeadj <- seasadj(fitd)
plot(eeadj)
```

### e)	Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier? 

```{r }
ukcars2 <- ts(c(ukcars[1:64],ukcars[65]+500,ukcars[66:113] ),start=c(1977,1),frequency=4)
fitd2 <- decompose(ukcars2)
plot(fitd2)
eeadj2 <- seasadj(fitd2)
plot(eeadj2)

```

The effect of this outlier is more local, the trend remains the same.

### f)	Does it make any difference if the outlier is near the end rather than in the middle of the time series? 
```{r }
ukcars3 <- ts(c(ukcars[1:110],ukcars[111]+500,ukcars[112:113] ),start=c(1977,1),frequency=4)
fitd3 <- decompose(ukcars3)
plot(fitd3)
eeadj3 <- seasadj(fitd3)
plot(eeadj3)
```

### g)	Use STL to decompose the series. 

```{r }
fit <- stl(ukcars, s.window=5)
plot(fit)

plot(ukcars, col="gray", main="UK cars" , ylab="New Cars", xlab= "")
lines(fit$time.series[,2],col="red" ,ylab="Trend")


```